\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-07-11 fully-abstract-comp/popl2019/paper.tex

\subsubsection{Worlds}
\label{subsec:worlds}
Generally speaking, a world is used to model the memory of a language in a logical relation.
\srccm{} has three kinds of memory, the heap, the call stack, and the free stack, and our worlds model this by being triples of sub-worlds.
All of the sub-worlds are partial maps from region names $\RegionName$ (modelled as natural numbers) to regions where the regions differ for each of the sub-worlds based on what they model.
Generally, the regions consist of a region type and a world indexed relation that specifies the memories allowed by the region.
The region types models linearity as they indicate how the memory can be addressed.
We have three types of regions: $\pure$, $\spatial$, and $\spatialo$.
The $\pure$ regions can be addressed by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ regions by nothing.
To prevent duplication of linear capabilities, two linear capabilities cannot address the same $\spatialo$, so ownership has to be distributed between worlds.
To this end, we have the relation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except for the ownership which has to be disjoint.
%We define $W_1 \uplus W_2$ which joins together two worlds with disjoint regions.
The unaddressable $\spatial$ region makes sure that the memory it represents is modelled.
Linear capabilities can be seen as revocable which is modelled by the $\revoked$ region.
Finally, the $\pure$ region also contains a seal interpretation function.
The seal interpretation is a world-indexed function that specifies the sealables a seal can be used to seal.
We split the regions in two categories:
\begin{align*}
  \Regionh &= 
             \left\{\begin{multlined}
               \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
               (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
             \end{multlined}\right.\\
  \Regions &={}
             \left\{\begin{multlined}
               \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup{} \\
               \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\cup
               \{\revoked\}
             \end{multlined}\right.
\end{align*}
and define the sub-worlds as follows:
\begin{align*}
  \Worldh  &={} \RegionName \parfun (\Regions + \Regionh)\\
  \Worlds  &={} \RegionName \parfun (\Regions \times \Addr)\\
  \Worldfs &={} \RegionName \parfun \Regions
\end{align*}
The world for the heap can contain every kind of region because it can be addressed by both normal and linear capabilities.
The call stack and free stack is only addressable by linear capabilities, so their sub-worlds only allow regions from $\Regions$.
Regions in $\Worlds$ correspond to a stack frame, so they are paired with an address which models the return address.

Define worlds as $\World = (\Worldh \times \Worlds \times \Worldfs)$. They are recursive as they contain world indexed relations. Recursive worlds are common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} to construct $\Wor$ which solves the circularity\footnote{We have cheated in the presentation and included the solution to the circularity already.}.
% \begin{lemma}
%   \label{thm:recursive-domain-eq}
%   There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e.\and there exists an isomorphism $\xi$ such that
%   \[
%     \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
%   \]
%   and for $\hat{W},\hat{W}' \in \Wor: \quad  \hat{W'} \future \hat{W}\text{ iff }\xi(\hat{W'}) \future \xi(\hat{W})$
% \end{lemma}
The future world relation $\future$ is fairly standard.
It is extensional and requires regions to stay the same type in future worlds except $\spatial$ regions can become $\spatialo$ or $\revoked$.

\subsubsection{Code region}
The code region is a central region to the logical relation that models the code memory of a component.
It is especially interesting because its seal interpretation function specifies what the return and closure seals in a component can seal.
The code region is defined as follows:
\[
  \codereg{\sigrets,\sigcloss,\mscode,\gc} \defeq (\pure,
  H^\mathrm{code,\square} \; \sigrets \; \sigcloss \; \mscode \; \gc,
  H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc)
\]
The code region is $\pure$ which makes it persistent in future worlds and reflects the fact that it is addressed by normal capabilities.
The memory relation $H^\mathrm{code,\square} \; (\dots)$ only relates $\mscode$ to itself which corresponds with our component setup where code memory is read-only.
$H^\mathrm{code,\square} \; (\dots)$ requires $\mscode$ to be well-formed and that its contents is non-linear and in the value relation.
As we will see later, we have two value relations, a trusted and an untrusted, and the value relation used depends on whether the code region governs trusted or untrusted code.
For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have, and the trusted value relation includes everything that is in the untrusted value relation, but also the things that are not safe to pass away in a call, e.g.\ return seals.

@@

The seal interpretation function $H^\mathrm{code,\square}_\sigma \; (\dots)$ dictates what the return seals $\sigrets$ and closure seals $\sigclos$ can seal.
The return seals can only be used to seal return pointers that corresponds to the return from an call in the code of the region.
Code return pointers must point to the first address after a call, and the call must be associated with the return seal in question.
The corresponding capability on \trgcm{} should point to the address just after the \texttt{xjmp} in the call code (figure~\ref{fig:call-code}) as it needs to perform the safety checks that are built into the call on \srccm{}.
The code return pointer corresponds to a normal capability which means that it can be reused but always to return from the same call.
Because of this, the code pointer is not associated with any specific call frame as it can be used by any data return pointer generated from the same call point.
Reuse of return pairs is prevented by the data return pointer being linear, so the seal interpretation function related data return pointers to linear capabilities.
It also requires the data return pointer to correspond to one of the stack frames on the call stack.
Which is expressed by requiring the existence of a spatially owned region in $\Worlds$ that governs memory corresponding to the return pointers range of authority.
It must be spatially owned, so it is permitted to address it.
A region in $\Worlds$ is also associated with a return address, and it should be the return address of the call that uses the return seal in question.
\begin{multline*}
  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \; \sigma \; \hat{W}= \\
  \begin{array}[t]{l}
\left\{
    \begin{array}{l}
      \npair{\left(\retptrc(\baddr,\eaddr,\aaddr+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr+\mathrm{xjmp\_off}+1)\right)} \mid \dots
    %     \sigrets \subseteq \gsigrets \tand \\
    %     \quad\dom(\mscode) \subseteq \ta \tand 
    %     \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
    %     \quad\aaddr = \aaddr' + \retoffset \tand \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand\\
    %     \quad \sigma = \sigma_b + \offsigma \in \sigrets \tand
    %     \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
    \end{array}
      \right\} \cup \\
\left\{
    \begin{array}{l}
      \npair{(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1))} \mid \dots
        % \sigrets \subseteq \gsigrets \tand 
        % \dom(\mscode) \subseteq \ta \tand\\
        % \quad \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]},\ms_S,\ms_T \ldotp\\
        % \qquad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
        % \qquad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
        % \qquad \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
        % \qquad \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
    \end{array}
    \right\} \\
  \end{array}\\
    \text{ if } \sigma \in \sigrets
\end{multline*}

@@

\subsubsection{Expression relations and friends}
The logical relation has two expression relations.
They both relate capabilities that can safely be used for execution in configurations that satisfy a given world.
The first one $\mathcal{E}$ is standard as logical relations for capability machines go: It takes related register files and related memories, plugs the capabilities in question into the program counter, and requires it to observably behave the same way.
In our case, the only observable thing on the capability machines is termination.
The observation relation $\mathcal{O}$ (Figure~\ref{fig:obs-rel}) defines when a configuration observably approximates another\footnote{While we present this as a logical relation, it is in fact two logical approximations. However, the only difference between the two approximations is in the observation approximation.}.
\begin{figure}[htb]
  \centering
  \begin{align*}
  \lro[\preceq,(\ta,\stkb,\_,\_)] ={}&\{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
  \forall i \leq n \ldotp
    \src{\Phi_S \sterm[i]{\ta,\stkb}} \Rightarrow \Phi_T \trg{\term} \}\\
  \lro[\succeq,(\ta,\stkb,\_,\_)] ={}& \{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
  \forall i \leq n \ldotp 
    \Phi_T \trg{\term[i]} \Rightarrow \src{\Phi_S \sterm{\ta,\stkb}} \}
\end{align*}
\caption{The observation approximations.}
\label{fig:obs-rel}
\end{figure}
With sealed capabilities in the mix, we also need to be able to say that \texttt{xjmp}s are safe, but a sealed code capability only needs to be safe to execute when it has access to the related data capability.
Further, \texttt{xjmp}s are used to invoke return pairs which on \srccm{} is somewhat involved.
This is handled uniformly by the \texttt{xjmp}-expression relation $\mathcal{E}_{\mathrm{xjmp}}$.
It relates two pairs of capabilities by plugging them into related register-files and memories using the $\xjmpres{}$ function and requiring that the result is in the observation relation.
\extend{Write that it is untrusted register files + intuition for this}
  % \begin{multline*}
  %   \lre(W)= \\
  %   \left\{ \begin{array}{l}
  %     \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} | \\
  %     \quad\forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
  %     \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand\\
  %     \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}} \tand\\
  %     \qquad\Phi_T = (\ms_T,\reg_T) \tand \Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}} \tand W \oplus W_\lrrs \oplus W_\lrm\\
  %     \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
  %   \end{array}
  %   \right\}
  % \end{multline*}
% \begin{multline*}
%   \lrexj(W) =
%   \\\left\{\begin{array}{l}
%       \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} | \\
%       \quad \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
%       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand \\
%       \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_T = (\ms_T,\reg_T) \tand W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
%       \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp \Phi_S' = \src{\var{xjmpRes}}(v_{c,S},v_{d,S},\Phi_S) \tand\\
%       \quad\qquad\Phi_T' = \xjmpres{v_{c,T},v_{d,T},\Phi_T}\tand\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
%     \end{array}
%     \right\}
% \end{multline*}
% register file relation

@@

\subsubsection{Value relations}
The purpose of the value relations is two fold: It relates the words on \srccm{} to the words they correspond to on \trgcm{}, and it defines what pairs of words are safe, i.e.\ are unable to break the guarantees of the capability machines including the LSE and WBCF guaranteed by \srccm{}.
As previously explained, we try to ensure LSE and WBCF for some trusted code that we expect to behave reasonable, e.g.\ by treating return seals in a reasonable way.
This means that it is safe to give return seals to trusted code; but at the same time, we have some code that we do not trust and that we want to guarantee LSE and WBCF against.
This means that there is a difference between what is safe to give to the trusted code and the untrusted code which we capture by having a trusted and untrusted value relation.
The two value relations are sketched in Figure~\ref{fig:value-relation}.

The untrusted value relation relates words that can safely be given to untrusted code.
Data is always safe, and a piece of data relates to itself.
A stack pointers is related to a linear capability with the same (non-executable) permission, range of authority, and current address.
The stack pointer should only be able to produce safe words, so a stack pointer with read permission can only produce safe words, i.e.\ words in $\mathcal{V}_\untrusted$, and a stack pointer with write permission can at least write safe words.
Regions model the contents of memory, so to determine safety of a stack pointer there must be an appropriate spatially owned region in the free stack part of the world.
Syntactically equal memory capabilities are related in a similar way to stack capabilities when they have read-write permission.
The difference is that memory capabilities point to the heap rather than the free stack, so the regions they depend on are picked from the heap sub-world.
An executable memory capability should always point to code which means that there should be a code region governing the memory it points to.
The trusted return seals and closure seals cannot be given safely to untrusted code, so a set of seal is only in the untrusted value relation when it is disjoint from those, and they may only be used to seal things in $\mathcal{V}_\untrusted$.
% The write condition is called $\stackWriteCond{}$
% \begin{multline*}
%   \stackWriteCond{W} =\\
%   \left\{
%     \begin{array}{l}\npair{A} \mid 
%       \exists S \subseteq \addressable{\linear, \pwfree} \ldotp  \exists R : S \fun \powerset{\nats} \\
%       \quad \biguplus_{r\in S} R(r) \supseteq A \tand \forall r \in S \ldotp |R(r)| = 1  \tand \\
%       \quad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \tand \pwfree(r) \text{ is address-stratified}
%     \end{array}
%   \right\}
% \end{multline*}
When a piece of untrusted code is called, it is passed a return pair consisting of two sealed capabilities.
Calling untrusted code should be safe, so sealed capabilities must be related in the untrusted value relation.
The seal interpretation function decides what a seal can be used for, so for a pair $(\src{\sealed{\sigma,\vsc_S}}, \sealed{\sigma,\vsc_T})$ to be in the untrusted value relation, $(\vsc_S,\vsc_T)$ must be in the seal interpretation function for $\sigma$.
Further, it should be safe to use the sealed capability, i.e.\ jump to it with \texttt{xjmp}, which is exactly what $\mathcal{E}_{\mathrm{xjmp}}$ expresses, so when paired up with another pair from the seal interpretation function for $\sigma$ the quadruple must be in $\mathcal{E}_{\mathrm{xjmp}}$.

@@

\subsubsection{Fundamental theorem}
With the logical relation defined, we state the fundamental theorem of logical relations (FTLR).
\begin{lemma}[FTLR]
  \label{thm:ftlr}
  For all $n,W,\lin,\baddr,\eaddr,\aaddr$,
  If $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
  and either $[\baddr,\eaddr] \subseteq \ta$ and $({((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps or
$[\baddr,\eaddr] \mathrel{\#} \ta$,
then 
  \[
    \npair{\left(((\rx,\normal),\baddr,\eaddr,\aaddr), ((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \in \lre(W) \qedhere
  \]
\end{lemma}
Roughly speaking, the FTLR says that any executable capability used for execution in a safe environment won't break the capability machine guarantees.
This is, however, only true for syntactically well-formed code, which the $\xReadCond{}$ entails, and for trusted code the semantic reasonability condition must also be satisfied.
The proof of the FTLR is done by a nested induction where each instruction that could be executed is considered.
The lemma is not directly used in the proof sketch we present in Section~\ref{subsec:proof-sketch}, but it is crucial to the proofs of the lemmas in that section.

@@

The next is a compatibility lemma for plugging that says that given a context and component each related to itself in the component relation, then plugging the context with the component will give an executable configuration in the executable configuration relation.

>>>>

\subsubsection{Worlds}
\label{subsec:worlds}
Generally speaking, a world is used to model the memory of a language in a logical relation.
\srccm{} has three kinds of memory: the heap, the call stack, and the free stack, and our worlds model this by being triples of sub-worlds.
All of the sub-worlds are partial maps from region names (modelled as natural numbers) to regions where the regions differ for each of the sub-worlds based on what they model.
Generally, the regions consist of a region type and a world indexed relation that specifies the memories allowed by the region.
The region types model linearity as they indicate how the memory can be addressed.
We have three types of regions: $\pure$, $\spatial$, and $\spatialo$.
The $\pure$ regions can be addressed by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ regions by nothing.
To prevent duplication of linear capabilities, two linear capabilities cannot address the same $\spatialo$, so ownership has to be distributed between worlds.
To this end, we have the relation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except for the ownership which has to be disjoint.
%We define $W_1 \uplus W_2$ which joins together two worlds with disjoint regions.
The unaddressable $\spatial$ region makes sure that the memory it represents is modelled.
Linear capabilities can be seen as revocable which is modelled by the $\revoked$ region.
Finally, the $\pure$ region also contains a seal interpretation function.
The seal interpretation is a world-indexed function that specifies the sealables a seal can be used to seal.
We split the regions in two categories:
\begin{align*}
  \Regionh &= 
             \left\{\begin{multlined}
               \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
               (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
             \end{multlined}\right.\\
  \Regions &={}
             \left\{\begin{multlined}
               \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup{} \\
               \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\cup
               \{\revoked\}
             \end{multlined}\right.
\end{align*}
and define the sub-worlds as follows:
\begin{align*}
  \Worldh  &={} \RegionName \parfun (\Regions + \Regionh)\\
  \Worlds  &={} \RegionName \parfun (\Regions \times \Addr)\\
  \Worldfs &={} \RegionName \parfun \Regions
\end{align*}
The world for the heap can contain every kind of region because it can be addressed by both normal and linear capabilities.
The call stack and free stack are only addressable by linear capabilities, so their sub-worlds only allow regions from $\Regions$.
Regions in $\Worlds$ correspond to a stack frame, so they are paired with an address which models the return address.

Define worlds as $\World = (\Worldh \times \Worlds \times \Worldfs)$. They are recursive as they contain world indexed relations. Recursive worlds are common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} to construct $\Wor$ which solves the circularity\footnote{We have cheated in the presentation and included the solution to the circularity already.}.
% \begin{lemma}
%   \label{thm:recursive-domain-eq}
%   There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e.\and there exists an isomorphism $\xi$ such that
%   \[
%     \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
%   \]
%   and for $\hat{W},\hat{W}' \in \Wor: \quad  \hat{W'} \future \hat{W}\text{ iff }\xi(\hat{W'}) \future \xi(\hat{W})$
% \end{lemma}
The future world relation $\future$ is fairly standard.
It is extensional and requires all but $\spatial$ regions to stay the same type in future worlds.
$\spatial$ regions can become $\spatialo$ or $\revoked$.

\subsubsection{Code region}
The code region is a central region to the logical relation that models the code memory of a component.
It is especially interesting because its seal interpretation function specifies what the return and closure seals in a component can seal.
The code region is defined as follows:
\[
  \codereg{\sigrets,\sigcloss,\mscode,\gc} \defeq (\pure,
  H^\mathrm{code,\square} \; \sigrets \; \sigcloss \; \mscode \; \gc,
  H^\mathrm{code}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \gc)
\]
The code region is $\pure$ which makes it persistent in future worlds and reflects the fact that it is addressed by normal capabilities.
The memory relation $H^\mathrm{code,\square} \; (\dots)$ only relates $\mscode$ to itself which agrees with our component setup where code memory is read-only.
$H^\mathrm{code,\square} \; (\dots)$ requires $\mscode$ to be well-formed and that its contents is non-linear and in the value relation.
As we will see later, we have two value relations: a trusted and an untrusted.
The value relation used depends on whether the code region governs trusted or untrusted code.
For now it suffices to know that the untrusted value relation contains everything that untrusted code can safely have, and the trusted value relation includes everything that is in the untrusted value relation in addition to the things that are not safe to pass away in a call, e.g.\ return seals.

@@

The seal interpretation function $H^\mathrm{code,\square}_\sigma \; (\dots)$ dictates what the return seals $\sigrets$ and closure seals $\sigclos$ can seal.
The return seals can only be used to seal return pointers that corresponds to the return from an call in the code of the region.
Code return pointers must point to the first address after a call, and the call must be associated with the return seal in question.
The corresponding capability on \trgcm{} should point to the address just after the \texttt{xjmp} in the call code (figure~\ref{fig:call-code}) as it needs to perform the safety checks that are built into the call on \srccm{}.
The code return pointer corresponds to a normal capability which means that it can be reused but always to return from the same call.
Because of this, the code pointer is not associated with any specific call frame as it can be used by any data return pointer generated from the same call point.
Reuse of return pairs is prevented by the data return pointer being linear, so the seal interpretation function related data return pointers to linear capabilities.
It also requires the data return pointer to correspond to one of the stack frames on the call stack.
This is expressed by requiring the existence of a spatially owned region in $\Worlds$ that governs memory corresponding to the return pointer's range of authority.
It must be spatially owned, so it is permitted to address it.
A region in $\Worlds$ is also associated with a return address, and it should be the return address of the call that uses the return seal in question.
\begin{multline*}
  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \; \sigma \; \hat{W}= \\
  \begin{array}[t]{l}
\left\{
    \begin{array}{l}
      \npair{\left(\retptrc(\baddr,\eaddr,\aaddr+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr+\mathrm{xjmp\_off}+1)\right)} \mid \dots
    %     \sigrets \subseteq \gsigrets \tand \\
    %     \quad\dom(\mscode) \subseteq \ta \tand 
    %     \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
    %     \quad\aaddr = \aaddr' + \retoffset \tand \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand\\
    %     \quad \sigma = \sigma_b + \offsigma \in \sigrets \tand
    %     \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
    \end{array}
      \right\} \cup \\
\left\{
    \begin{array}{l}
      \npair{(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1))} \mid \dots
        % \sigrets \subseteq \gsigrets \tand 
        % \dom(\mscode) \subseteq \ta \tand\\
        % \quad \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]},\ms_S,\ms_T \ldotp\\
        % \qquad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
        % \qquad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
        % \qquad \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
        % \qquad \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
    \end{array}
    \right\} \\
  \end{array}\\
    \text{ if } \sigma \in \sigrets
\end{multline*}

@@

\subsubsection{Expression relations and friends}
The logical relation has two expression relations.
They both relate capabilities that can safely be used for execution in configurations that satisfy a given world.
The first one, $\mathcal{E}$, is standard as logical relations for capability machines go: It takes related register files and related memories, plugs the capabilities in question into the program counter, and requires it to observably behave the same way.
In our case, the only observable thing on the capability machines is termination.
The observation relation $\mathcal{O}$ (Figure~\ref{fig:obs-rel}) defines when a configuration observably approximates another\footnote{While we present this as a logical relation, it is in fact two logical approximations. However, the only difference between the two approximations is in the observation approximation.}.
\begin{figure}[htb]
  \centering
  \begin{align*}
  \lro[\preceq,(\ta,\stkb,\_,\_)] ={}&\{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
  \forall i \leq n \ldotp
    \src{\Phi_S \sterm[i]{\ta,\stkb}} \Rightarrow \Phi_T \trg{\term} \}\\
  \lro[\succeq,(\ta,\stkb,\_,\_)] ={}& \{ \npair{\left(\src{\Phi_S},\Phi_T\right)} \mid
  \forall i \leq n \ldotp 
    \Phi_T \trg{\term[i]} \Rightarrow \src{\Phi_S \sterm{\ta,\stkb}} \}
\end{align*}
\caption{The observation approximations.}
\label{fig:obs-rel}
\end{figure}
With sealed capabilities in the mix, we also need to be able to say that \texttt{xjmp}s are safe, but a sealed code capability only needs to be safe to execute when it has access to the related data capability.
Further, \texttt{xjmp}s are used to invoke return pairs which on \srccm{} is somewhat involved.
This is handled uniformly by the \texttt{xjmp}-expression relation $\mathcal{E}_{\mathrm{xjmp}}$.
It relates two pairs of capabilities by plugging them into related register-files and memories using the $\xjmpres{}$ function and requiring that the result is in the observation relation.
\extend{Write that it is untrusted register files + intuition for this}
  % \begin{multline*}
  %   \lre(W)= \\
  %   \left\{ \begin{array}{l}
  %     \npair{\stpair[.]{v_{c,S}}{v_{c,T}}} | \\
  %     \quad\forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
  %     \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand\\
  %     \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_S' = \Phi_S \updReg{\pcreg}{v_{c,S}} \tand\\
  %     \qquad\Phi_T = (\ms_T,\reg_T) \tand \Phi_T' = \Phi_T\updReg{\pcreg}{v_{c,T}} \tand W \oplus W_\lrrs \oplus W_\lrm\\
  %     \qquad\Rightarrow\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
  %   \end{array}
  %   \right\}
  % \end{multline*}
% \begin{multline*}
%   \lrexj(W) =
%   \\\left\{\begin{array}{l}
%       \npair{\stpair[.]{v_{c,S},v_{d,S}}{v_{c,T},v_{d,T}}} | \\
%       \quad \forall n' \leq n, \src{\reg_S}, \reg_T, \src{\ms_S}, \ms_T, \src{\ms_\stk}, \src{\stk}, W_\lrrs , W_\lrm \ldotp \\
%       \qquad\npair[n']{\stpair{\reg}{\reg}} \in \lrr(\{\rdata\}) (W_\lrrs ) \tand \memSat[n']{\stpair[.]{\ms_S,\stk,\ms_\stk}{\ms_T}}{W_\lrm} \tand \\
%       \qquad\Phi_S = \src{(\ms_S,\reg_S,\stk, \ms_\stk)} \tand \Phi_T = (\ms_T,\reg_T) \tand W \oplus W_\lrrs \oplus W_\lrm \text{ is defined }\\
%       \qquad\Rightarrow \exists \Phi_S',\Phi_T'\ldotp \Phi_S' = \src{\var{xjmpRes}}(v_{c,S},v_{d,S},\Phi_S) \tand\\
%       \quad\qquad\Phi_T' = \xjmpres{v_{c,T},v_{d,T},\Phi_T}\tand\npair[n']{\left(\Phi_S', \Phi_T' \right)}\in \lro
%     \end{array}
%     \right\}
% \end{multline*}
% register file relation

@@

\subsubsection{Value relations}
The purpose of the value relations is twofold: It relates the words on \srccm{} to the words they correspond to on \trgcm{}, and it defines what pairs of words are safe, i.e.\ are unable to break the guarantees of the capability machines including the LSE and WBCF guaranteed by \srccm{}.
As previously explained, we try to ensure LSE and WBCF for some trusted code that we expect to behave well, e.g.\ by treating return seals in a reasonable way.
This means that it is safe to give return seals to trusted code; but at the same time, we have some code that we do not trust and that we want to guarantee LSE and WBCF against.
This means that there is a difference between what is safe to give to the trusted code and the untrusted code which we capture by having a trusted and untrusted value relation.
The two value relations are sketched in Figure~\ref{fig:value-relation}.

The untrusted value relation relates words that can safely be given to untrusted code.
Data is always safe, and a piece of data relates to itself.
A stack pointer is related to a linear capability with the same (non-executable) permission, range of authority, and current address.
The stack pointer should only be able to produce safe words, so a stack pointer with read permission can only produce safe words, i.e.\ words in $\mathcal{V}_\untrusted$, and a stack pointer with write permission can at least write safe words.
Regions model the contents of memory, so to determine safety of a stack pointer there must be an appropriate spatially owned region in the free stack part of the world.
Syntactically equal memory capabilities are related in a similar way to stack capabilities when they have read-write permission.
The difference is that memory capabilities point to the heap rather than the free stack, so the regions they depend on are picked from the heap sub-world.
An executable memory capability should always point to code which means that there should be a code region governing the memory it points to.
The trusted return seals and closure seals cannot be given safely to untrusted code, so a set of seals is only in the untrusted value relation when it is disjoint from those, and they may only be used to seal things in $\mathcal{V}_\untrusted$.
% The write condition is called $\stackWriteCond{}$
% \begin{multline*}
%   \stackWriteCond{W} =\\
%   \left\{
%     \begin{array}{l}\npair{A} \mid 
%       \exists S \subseteq \addressable{\linear, \pwfree} \ldotp  \exists R : S \fun \powerset{\nats} \\
%       \quad \biguplus_{r\in S} R(r) \supseteq A \tand \forall r \in S \ldotp |R(r)| = 1  \tand \\
%       \quad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \tand \pwfree(r) \text{ is address-stratified}
%     \end{array}
%   \right\}
% \end{multline*}
When a piece of untrusted code is called, it is passed a return pair consisting of two sealed capabilities.
Calling untrusted code should be safe, so sealed capabilities must be related in the untrusted value relation.
The seal interpretation function decides what a seal can be used for, so for a pair $(\src{\sealed{\sigma,\vsc_S}}, \sealed{\sigma,\vsc_T})$ to be in the untrusted value relation, $(\vsc_S,\vsc_T)$ must be in the seal interpretation function for $\sigma$.
Further, it should be safe to use the sealed capability, i.e.\ jump to it with \texttt{xjmp}, which is exactly what $\mathcal{E}_{\mathrm{xjmp}}$ expresses, so when paired up with another pair from the seal interpretation function for $\sigma$ the quadruple must be in $\mathcal{E}_{\mathrm{xjmp}}$.

@@

\subsubsection{Fundamental theorem}
With the logical relation defined, we state the fundamental theorem of logical relations (FTLR).
\begin{lemma}[FTLR]
  \label{thm:ftlr}
  For all $n,W,\lin,\baddr,\eaddr,\aaddr$,
  if\ $\npair{[\baddr,\eaddr]} \in \xReadCond{W}$
  and either $[\baddr,\eaddr] \subseteq \ta$ and $({((\rx,\normal),\baddr,\eaddr,\aaddr)}$ behaves reasonably up to $n$ steps or
$[\baddr,\eaddr] \mathrel{\#} \ta$,
then 
  \[
    \npair{\left(((\rx,\normal),\baddr,\eaddr,\aaddr), ((\rx,\normal),\baddr,\eaddr,\aaddr)\right)} \in \lre(W) \qedhere
  \]
\end{lemma}
Roughly speaking, the FTLR says that any executable capability used for execution in a safe environment won't break the capability machine guarantees.
This is, however, only true for syntactically well-formed code, which the $\xReadCond{}$ entails, and for trusted code the semantic reasonability condition must also be satisfied.
The proof of the FTLR is done by a nested induction where each instruction that could be executed is considered.
The lemma is not directly used in the proof sketch we present in Section~\ref{subsec:proof-sketch}, but it is crucial to the proofs of the lemmas in that section.

@@

The next is a compatibility lemma for plugging that says: given a context and component each related to itself in the component relation, plugging the context with the component will give an executable configuration in the executable configuration relation.
